---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  message = FALSE
)
```
# budyko
The goal of `budyko` is to facilitate fitting of hydrologic data to Budyko-type curves. The package includes the original Budyko (1974) formulation, as well as several single-parameter formulations like those of Fu (1981), Turc (1954), Pike (1964), and Wang & Tang (2014). 

## Installation

You can install budyko from github using devtools:

```{r gh-installation, eval = FALSE}
devtools::install_github("tylerbhampton/budykoR")
```

## Sample Plots and Data

We can use the function `blankBC` to create a blank Budyko Curve plot, which includes the energy and water boundaries.

```{r}
library(ggplot2)
library(budyko)
library(wesanderson)

blankBC
blankBC+coord_cartesian(xlim=c(0,3))
```

The package comes with a sample dataset. This is annual Water-Year (October through September) data for 30 years. blankBC has pre-set aesthetics for columns `PET.P` on the x axis and `AET.P` on the y axis. We can simply add `testdata` to the plot.

```{r}
testdata=budyko::testdata
head(testdata)
blankBC+geom_point(data=testdata)+coord_cartesian(xlim=c(0,5))
```

## Curves and Curve Fitting

The package contains several Budyko-type curves. We can use the default setting of `budyko_sim()` to generate the original Budyko curve.

```{r}
ogbudyko=budyko_sim()
blankBC+geom_line(data=ogbudyko)+coord_cartesian(xlim=c(0,5))
blankBC+
  geom_line(data=ogbudyko)+
  geom_point(data=testdata)+
  coord_cartesian(xlim=c(0,5))
```

We can also use `budyko_fit()` to custom fit the data.

```{r}
fit1=budyko_fit(data=testdata,method="Fu",dif="mae",silent = TRUE)
fit1
sim1=budyko_sim(fit=fit1)
blankBC+
  geom_line(data=ogbudyko,col=2)+
  geom_line(data=sim1)+
  geom_point(data=testdata)+
  coord_cartesian(xlim=c(0,5))
```


```{r}
#load data
datos <- read.csv("~/R/Budyko/data/año.csv")
#set column name to PET.P needed
names(datos)[6]<-"PET.P"

#fit budyko curve to data
#set the method and create curve according data dif = mae or nls, rsq, metthod is Fu, Turc-Pike, Wang-Tang
fit1=budyko_fit(data=datos,method="fu",dif="nls",silent = TRUE)

#simulate budyko curve
sim1=budyko_sim(fit = fit1)

#Create plot with info above

blankBC+geom_line(data=sim1, col=2)+
  theme_classic()+
  geom_point(data=datos, aes(y=AET.P, x=PET.P, color = factor(year)), size = 3, shape=17)+
  coord_cartesian(xlim=c(0,5))+
  scale_color_manual(values=wes_palette(n=4, name="Moonrise2"))+
  theme(legend.title = element_blank(), legend.position = "bottom")
  
```
